---
title: "1_mw_sleep"
author: "Jackie"
date: "02/25/22"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

This script wrangles MW subjctive sleep data for further analysis

# Loading libraries

```{r libraries, message=FALSE}
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(lubridate)
library(lmerTest)
library(psych)
```

# Read in data

## MW sleep data

> Reading in usable sleep data (>=3 days of data)

```{r read mw sleep, message=FALSE}
sleep_mw_usable_fp <- "~/Dropbox/JSK_SC/EMA/data_3_days_usability/2_daily_sleep_usable_data.csv"

sleep_mw_usable <-
  read_csv(sleep_mw_usable_fp) %>%
  mutate(
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = as.character(ELS_ID),
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  dplyr::select(-c(email, device_ID))
```
## Summary sheet

```{r}
sum_sheet <- "~/Dropbox/JSK_SC/curating_data/3_actigraphy_ema_data_summary.csv"
sum_sheet_df <- 
  read_csv(sum_sheet) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  ) %>%
  dplyr::select(ELS_ID, simiul_ahc_t3, simul_ahc_t4, starts_with("use")) %>%
  mutate(
   simul =
      ifelse(simiul_ahc_t3 == "yes" | simul_ahc_t4 == "yes", "1", "0")
  ) %>%
  filter(simul == "1")
```
## MERGING
```{r}
 sleep_merge <-
  dplyr::left_join(
    sleep_mw_usable,
    sum_sheet_df,
    by = "ELS_ID"
  )

sleep_merge2 <-
  sleep_merge %>%
  filter(simul == "1")

sleep_merge2 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::summarise(
    n = n()
  )
```

### cleaning variable names

```{r}
sleep_mw_usable_clean <-
  sleep_merge2 %>%
  rename(
    DailySleep_hrs = 
      `How many hours did you sleep last night?`,
    DailySleep_satisfaction = 
      `How restful or satisfying was your sleep last night?`
  ) %>%
  mutate(
    DailySleep_hrs_rec =
      recode_factor(
        DailySleep_hrs,
        "5hr or less" = 5,
        "6 hr." = 6,
        "7hr." = 7,
        "8hr." = 8,
        "9hr. or more" = 9
        )
    ) %>%
  mutate(
    DailySleep_hrs_rec = 
      as.character(DailySleep_hrs_rec)
    ) %>%
  mutate(
    DailySleep_hrs_rec =
      as.numeric(DailySleep_hrs_rec)
    )
```

### parsing dates

```{r parsing dates}
sleep_mw_usable_clean <-
  sleep_mw_usable_clean %>%
  mutate(
    sleep_trigger_date = dmy(sleep_trigger_date),
    sleep_survey_start_date = dmy(sleep_survey_start_date)
  )
```

### creating day and week variables

```{r}
# creating day and week variables
sleep_mw_usable_clean_day <-
  sleep_mw_usable_clean %>%
  dplyr::mutate(
    day = day(sleep_trigger_date),
    wday = lubridate::wday(sleep_trigger_date, label = TRUE),
    ELS_ID = factor(ELS_ID)
  ) 

# day num
sleep_mw_usable_clean_dayorder  <-
  sleep_mw_usable_clean_day %>%
  group_by(ELS_ID) %>% 
  mutate(
    respnumber = order(sleep_trigger_date)
    )
```

### wkday or wkend

```{r}
sleep_mw_usable_clean_dayorder2 <-
  sleep_mw_usable_clean_dayorder %>%
  mutate(
    week =
      ifelse(
        wday == "Sun" |
          wday == "Sat",
        "wkend",
        "wkday"
      )
  ) 
```



## merging with  summary sheet
```{r}
sleep_merge3 <-
  sleep_mw_usable_clean_dayorder2 %>%
  mutate(
    use_ema_date_t3 = mdy(use_ema_date_t3),
    use_ema_date_t4 = mdy(use_ema_date_t4),
    num_of_days_btwn_t3 =
      abs((use_ema_date_t3 %--% sleep_survey_start_date)/ddays(1)),
    num_of_days_btwn_t4 =
      abs((use_ema_date_t4 %--% sleep_survey_start_date)/ddays(1)),
    num_of_days_btwn =
      ifelse(
        is.na(num_of_days_btwn_t3),
        num_of_days_btwn_t4,
        num_of_days_btwn_t3
      )
    )

greaterthan14 <- 
  sleep_merge3 %>%
  filter(
    num_of_days_btwn > 14
  )

sleep_merge4 <-
  sleep_merge3 %>%
  filter(
    num_of_days_btwn < 60
  ) %>% # now max is 14 days
  dplyr::select(-c(num_of_days_btwn_t3, num_of_days_btwn_t4, starts_with("use"), simiul_ahc_t3, simul_ahc_t4))

sleep_merge5 <-
  sleep_merge4 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::mutate(
    sleep_num_resp = n(),
    sleep_rr = sleep_num_resp/14
  )

```

### summmary

```{r}
sleep_merge5 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::summarise(
    n = n()
  ) # 100 subjects

write_csv(sleep_merge5, "simul_ema_subjs.csv")
```

# Data viz

> for those with simulataneous ema and actigraphy

```{r viz sleep sat and hours}
sleep_merge5 %>%
  ggplot(
    aes(respnumber, scale(DailySleep_satisfaction), group = ELS_ID)
    ) +
  geom_line(color = "blue") +
  geom_line(aes(x = respnumber, y = scale(DailySleep_hrs_rec), group = ELS_ID), color = "blue", alpha = .5) +
  facet_wrap(~ELS_ID) +
  scale_x_continuous(limits = c(1,15)) +
  theme_classic() +
  theme(
  strip.text.x = element_blank(),
  axis.title.y = element_text(size = 10)
  ) +
  labs(
    x = "Day",
    y = "Sleep Satisfaction (dark blue), Sleep Duration (light blue)",
    title = "Trajectories of Self-Reported Sleep Satisfaction and Duration"
    )

ggsave("daily_mw_sleep_traj.png", width = 7, height = 5.5, dpi = 400)
```


### distribution of weekday vs. weekend and association with sleep sat and hrs

```{r}
level_key <- c(wkday = "weekday", wkend = "weekend")
sleep_merge6 <-
  sleep_merge5 %>%
  mutate(
    week = factor(week),
    week =  recode_factor(week,!!!level_key),
    dayorder = as.numeric(respnumber),
    DailySleep_hrs_rec = as.numeric(DailySleep_hrs_rec)
  )

# summary
sleep_merge6 %>%
  dplyr::group_by(ELS_ID, week) %>%
  dplyr::summarise(
    n = n()
  )


sleep_merge6 %>%
  drop_na(DailySleep_satisfaction) %>%
  drop_na(DailySleep_hrs_rec) %>%
  drop_na(week) %>%
  ggplot(
    aes(
      x = respnumber, 
      y = scale(DailySleep_satisfaction)
    )
  ) +
  geom_line(aes(group = ELS_ID), color = "blue", alpha = .5) +
  geom_smooth(se=FALSE, size = 2, colour = "darkblue") +
  geom_line(aes(x = respnumber, y = scale(DailySleep_hrs_rec), group = ELS_ID), color = "darkgoldenrod2", alpha = .5) +
  geom_smooth(aes(x = respnumber, y = scale(DailySleep_hrs_rec)), size = 2, se = FALSE, colour = "darkgoldenrod2") +
  facet_wrap(~week) +
  theme_classic() +
  scale_x_continuous(
    limits = c(1, 14),                    
    breaks = seq(1, 14, 1)
    ) +
  labs(
    x = "Day",
    y = "Sleep Satisfaction = blue, Sleep Duration = gold"
  )
ggsave("sleepsat_and_hours_byDay_byWkdayWkend.png", width = 7, height = 5)

```

#### tests
```{r}
# sleep satisfaction differences by weekday/wkend 
summary(lmer(scale(DailySleep_satisfaction) ~ scale(respnumber) +factor(week) + (1 + scale(dayorder)|ELS_ID), data = sleep_merge6))
## marginal effect (higher on wkends compared to wkdays)

# sleep duration differences by weekday/wkend 
summary(lmer(scale(DailySleep_hrs_rec) ~ scale(respnumber) + factor(week) + (1 + scale(respnumber)|ELS_ID), data = sleep_merge6))
##  marginal effect (higher on wkends compared to wkdays)
```

### write

```{r}
write_csv(sleep_merge5, "~/Dropbox/JSK_SC/1_sleep_puberty/_scripts/2_at-home_wrangling/mw_daily_sleep_longform.csv")
```

### Averaging over the two weeks to obtain representative estimates for these constructs


```{r}
mw_sleep_red <-
  sleep_merge6 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::mutate(
    dailysleep_sat_mean = mean(DailySleep_satisfaction, na.rm = TRUE),
    dailysleep_hrs_mean = mean(DailySleep_hrs_rec, na.rm = TRUE),
  ) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  dplyr::select(ELS_ID, device_OS, sleep_num_resp, sleep_rr, dailysleep_sat_mean, dailysleep_hrs_mean) %>%
  distinct() %>%
  ungroup()
```

### write

```{r}
write_csv(mw_sleep_red, "~/Dropbox/JSK_SC/1_sleep_puberty/_scripts/2_at-home_wrangling/EMA_daily_sleep_data_reduced.csv")
```

