---
title: "actigraphy"
author: "Jackie"
date: "03/25/22"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

This script wrangles actigraphy data for subsequent analyses

# Loading libraries

```{r libraries, message=FALSE}
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(lubridate)
library(lmerTest)
library(psych)
```

# Read in data

## Actigraphy sleep data

> Reading in usable sleep data (>=3 days of data)

```{r read act sleep, message=FALSE}
sleep_act_usable_fp <- "~/Dropbox/JSK_SC/actigraphy/1_usable_actigraphy_data.csv"

sleep_act_usable <-
  read_csv(sleep_act_usable_fp) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  ) %>%
  dplyr::select(-c(analysis_name))
```

## Summary sheet

```{r}
sum_sheet <- "~/Dropbox/JSK_SC/curating_data/3_actigraphy_ema_data_summary.csv"
sum_sheet_df <- 
  read_csv(sum_sheet) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  ) %>%
  dplyr::select(ELS_ID, simiul_ahc_t3, simul_ahc_t4, starts_with("use")) %>%
  mutate(
   simul =
      ifelse(simiul_ahc_t3 == "yes" | simul_ahc_t4 == "yes", "1", "0")
  ) %>%
  filter(simul == "1")

```
## MERGING
```{r}
sleep_act_usable2 <-
  sleep_act_usable %>%
  dplyr::mutate(
    ELS_ID = as.character(ELS_ID),
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = as.factor(ELS_ID)
  )
  

sleep_merge <-
  dplyr::left_join(
    sleep_act_usable2,
    sum_sheet_df,
    by = "ELS_ID"
  )

sleep_merge2 <-
  sleep_merge %>%
  filter(simul == "1")

sleep_merge2 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::summarise(
    n = n()
  )
```

### cleaning variable names

> filter only for sleep intervals, and selecting only efficiency, duration, and sleep_time
> filtering for valid sleep intervals

```{r, cleaning}
d_vars <- c("data_start_date", "efficiency","sleep_time")

sleep_act_usable_clean <-
  sleep_merge2 %>%
  filter(interval_type == "SLEEP") %>%  
  dplyr::select(
    ELS_ID, 
    Subject_ID_Timepoint,
    act_timepoint,
    use_actigraphy_date_t3,
    use_actigraphy_date_t4,
    interval_type,
    interval_number,
    data_start_date,
    start_date, # night of administration
    start_time,
    end_date,
    end_time,
    efficiency,
    sleep_time,
    inv_time_sw
    ) %>%
  filter(
    inv_time_sw == 0
    ) %>%
  mutate(
    sleep_time_hrs = sleep_time/60
  ) 


### 171 was not a simultaneous person
sleep_act_usable_clean2 <-
  sleep_act_usable_clean %>%
  mutate(
    drop =
      ifelse(
        str_detect(Subject_ID_Timepoint, "196-T3"), # keeping 196x
        "drop",
        NA_real_
      ),
    drop = 
      ifelse(
        str_detect(Subject_ID_Timepoint, "085-T3") &
          str_detect(data_start_date, "7/19/19"),
        "drop",
        drop
      )
    ) %>%
  filter(
    is.na(drop)
  ) %>%
  dplyr::select(-drop, -inv_time_sw) %>%
  distinct()

sleep_act_usable_clean2 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::summarise(
    n = n()
  )
```

### parsing dates

```{r parsing dates}
sleep_act_usable_clean3 <-
  sleep_act_usable_clean2 %>%
  mutate(
    act_trigger_date = mdy(start_date),
    act_start_date = mdy(data_start_date),
    act_end_date = mdy(end_date)
  ) %>%
  dplyr::select(-c(start_date, data_start_date, end_date, interval_type))
```

### creating day and week variables

```{r}
# creating day and week variables
sleep_act_usable_clean4 <-
  sleep_act_usable_clean3 %>%
  dplyr::mutate(
    day = day(act_trigger_date),
    wday = lubridate::wday(act_trigger_date, label = TRUE),
    ELS_ID = factor(ELS_ID)
  ) 

# day num
sleep_act_usable_clean5  <-
  sleep_act_usable_clean4 %>%
  group_by(ELS_ID) %>% 
  arrange(act_trigger_date, .by_group = TRUE) %>%
  mutate(
    respnumber = order(act_trigger_date)
    )
```

### wkday or wkend

```{r}
sleep_act_usable_clean6 <-
  sleep_act_usable_clean5 %>%
  mutate(
    week =
      ifelse(
        wday == "Sun" |
          wday == "Sat",
        "wkend",
        "wkday"
      )
  ) %>%
  relocate(respnumber, .after = interval_number) %>%
  mutate(
    match = interval_number == respnumber # everything matches
  ) %>%
  dplyr::select(-interval_number, -match)
```

## merging with  summary sheet
```{r}

sleep_act_usable_clean7 <-
  sleep_act_usable_clean6 %>%
  mutate(
    use_actigraphy_date_t3 = mdy(use_actigraphy_date_t3),
    use_actigraphy_date_t4 = mdy(use_actigraphy_date_t4),
    num_of_days_btwn_t3 =
      abs((use_actigraphy_date_t3 %--% act_start_date)/ddays(1)),
    num_of_days_btwn_t4 =
      abs((use_actigraphy_date_t4 %--% act_start_date)/ddays(1)),
    num_of_days_btwn =
      ifelse(
        is.na(num_of_days_btwn_t3),
        num_of_days_btwn_t4,
        num_of_days_btwn_t3
      )
    ) # number of days between the act start date in the data and the summary sheet overlap

sleep_act_usable_clean8 <-
  sleep_act_usable_clean7 %>%
  dplyr::select(-c(num_of_days_btwn_t3, num_of_days_btwn_t4, starts_with("use")))

sleep_act_usable_clean9 <-
  sleep_act_usable_clean8 %>%
  group_by(ELS_ID) %>%
  mutate(
    act_num_resp = n(),
    act_rr = act_num_resp/14
  )
```

### summmary

```{r}
sleep_act_usable_clean9 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::summarise(
    n = n()
  ) # 103 subjects
write_csv(sleep_act_usable_clean9, "simul_actigraphy_subjs.csv")
```

# Data viz

> for those with simulataneous ema and actigraphy

```{r viz sleep sat and hours}
sleep_act_usable_clean9 %>%
  ggplot(
    aes(respnumber, scale(efficiency), group = ELS_ID)
    ) +
  geom_line(color = "blue") +
  geom_line(aes(x = respnumber, y = scale(sleep_time_hrs), group = ELS_ID), color = "blue", alpha = .5) +
  facet_wrap(~ELS_ID) +
  scale_x_continuous(limits = c(1,15)) +
  theme_classic() +
  theme(
  strip.text.x = element_blank(),
  axis.title.y = element_text(size = 10)
  ) +
  labs(
    x = "Day",
    y = "Sleep Efficiency (dark blue), Sleep Duration (light blue)",
    title = "Trajectories of Objective Sleep Efficiency and Duration"
    )

ggsave("daily_act_sleep_traj.png", width = 7, height = 5.5, dpi = 400)
```


### distribution of weekday vs. weekend and association with sleep sat and hrs

```{r}
level_key <- c(wkday = "weekday", wkend = "weekend")
sleep_act_usable_clean10 <-
  sleep_act_usable_clean9 %>%
  mutate(
    week = factor(week),
    week =  recode_factor(week,!!!level_key),
    respnumber = as.numeric(respnumber)
  )

# summary
sleep_act_usable_clean10 %>%
  group_by(ELS_ID, week) %>%
  summarise(
    n = n()
  )


sleep_act_usable_clean10 %>%
  drop_na(efficiency) %>%
  drop_na(sleep_time_hrs) %>%
  drop_na(week) %>%
  ggplot(
    aes(
      x = respnumber, 
      y = scale(efficiency)
    )
  ) +
  geom_line(aes(group = ELS_ID), color = "blue", alpha = .5) +
  geom_smooth(se=FALSE, size = 2, colour = "darkblue") +
  geom_line(aes(x = respnumber, y = scale(sleep_time_hrs), group = ELS_ID), color = "darkgoldenrod2", alpha = .5) +
  geom_smooth(aes(x = respnumber, y = scale(sleep_time_hrs)), size = 2, se = FALSE, colour = "darkgoldenrod2") +
  facet_wrap(~week) +
  theme_classic() +
  scale_x_continuous(
    limits = c(1, 14),                    
    breaks = seq(1, 14, 1)
    ) +
  labs(
    x = "Day",
    y = "Sleep Efficiency = blue, Sleep Duration = gold"
  )
ggsave("sleepeff_and_hours_byDay_byWkdayWkend.png", width = 7, height = 5)

```

#### tests
```{r}
# sleep efficiency differences by weekday/wkend 
summary(lmer(scale(efficiency) ~ scale(respnumber) +factor(week) + (1 + scale(respnumber)|ELS_ID), data = sleep_act_usable_clean10))
## no effect of weekday/weekend

# sleep duration differences by weekday/wkend 
summary(lmer(scale(sleep_time_hrs) ~ scale(respnumber) + factor(week) + (1 |ELS_ID), data = sleep_act_usable_clean10))
## sig effect (more hours on weekend)
```

### write

```{r}
write_csv(sleep_act_usable_clean10, "~/Dropbox/JSK_SC/_scripts/2_at-home_wrangling/act_daily_sleep_longform.csv")
```

### Averaging over the two weeks to obtain representative estimates for these constructs


```{r}
act_sleep_red <-
  sleep_act_usable_clean10 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::mutate(
    efficiency_mean = mean(efficiency, na.rm = TRUE),
    sleep_time_hrs_mean = mean(sleep_time_hrs, na.rm = TRUE),
  ) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  dplyr::select(ELS_ID, act_num_resp, act_rr, efficiency_mean, sleep_time_hrs_mean) %>%
  distinct() %>%
  ungroup()
```

### write

```{r}
write_csv(act_sleep_red, "~/Dropbox/JSK_SC/_scripts/2_at-home_wrangling/ACT_daily_sleep_data_reduced.csv")
```