---
title: "3_usable_ema_emo_data"
author: "Jackie"
date: "04/13/22"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

This script wrangles MW subjctive emo data for further analysis

# Loading libraries

```{r libraries, message=FALSE}
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(lubridate)
library(lmerTest)
library(psych)
```

# Read in data

## MW EMO data

> Reading in usable sleep data (>=3 days of data)

```{r read mw sleep, message=FALSE}
emo_mw_usable_fp <- "~/Dropbox/JSK_SC/EMA/2_daily_emo_usable_data.csv"

emo_mw_usable <-
  read_csv(emo_mw_usable_fp) %>%
  mutate(
    ELS_ID = as.numeric(ELS_ID),
    ELS_ID = as.character(ELS_ID),
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  dplyr::select(-c(email, device_ID))
```
## Summary sheet

```{r}
sum_sheet <- "~/Dropbox/JSK_SC/curating_data/3_actigraphy_ema_data_summary.csv"
sum_sheet_df <- 
  read_csv(sum_sheet) %>%
  mutate(
    ELS_ID = factor(ELS_ID)
  ) %>%
  dplyr::select(ELS_ID, simiul_ahc_t3, simul_ahc_t4, starts_with("use")) %>%
  mutate(
   simul =
      ifelse(simiul_ahc_t3 == "yes" | simul_ahc_t4 == "yes", "1", "0")
  ) %>%
  filter(simul == "1")
```

### cleaning variable names

```{r}
emo_mw_usable_clean <-
  emo_mw_usable %>%
  dplyr::rename(
    Daily_Sad = 
      `Since the last prompt, please indicate the extent to which you felt sad`,
    Daily_Annoyed.Angry = 
      `Since the last prompt, please indicate the extent to which you felt annoyed/angry`,
    Daily_Grouchy.Cranky =
      `Since the last prompt, please indicate the extent to which you felt grouchy/cranky`,
    Daily_Worried =
      `Since the last prompt, please indicate the extent to which you felt worried`,
    Daily_Anxious =
      `Since the last prompt, please indicate the extent to which you felt anxious`,
    Daily_Happy =
      `Since the last prompt, please indicate the extent to which you felt happy`,
    Daily_Cheerful =
      `Since the last prompt, please indicate the extent to which you felt cheerful`,
    Daily_Excited =
      `Since the last prompt, please indicate the extent to which you felt excited`,
    Daily_Energetic =
      `Since the last prompt, please indicate the extent to which you felt energetic`,
    Daily_Bored =
      `Since the last prompt, please indicate the extent to which you felt bored`,
    Daily_Want_Others =
      `Since the last prompt, please indicate the extent to which you want to be with other people`,
    Daily_Want_Alone =
      `Since the last prompt, please indicate the extent to which you want to be alone`,
    Social_Context =
      `Right now, who are you with?`,
    Face_to_Face_Convo =
      `Right now, are you having a face-to-face conversation?`,
    Digital_Convo =
      `Right now, are you having a real-time digital (phone, text, Facebook, video) conversation?`
  ) %>%
  mutate(
    Social_Context = factor(Social_Context),
    Face_to_Face_Convo = factor(Face_to_Face_Convo),
    Digital_Convo = factor(Digital_Convo),
    Social_Context_rec =
      recode_factor(
        Social_Context,
        "alone" = "alone_only",
        "alone,strangers" = "alone_or_unknown_others",
        "strangers,alone" = "alone_or_unknown_others",
        "alone,acquaintance(s)" = "alone_or_semiknown_others",
        "acquaintance(s),alone" = "alone_or_semiknown_others",
        "acquaintance(s),strangers" = "semiknown_or_unknown_others",
        "strangers,acquaintance(s)" = "semiknown_or_unknown_others",
        "acquaintance(s)" = "semiknown_others",
        "strangers" = "unknown_others"
        ),
    Social_Context_rec = as.character(Social_Context_rec)
  ) %>%
  mutate(
    Social_Context_rec2 = NA,
    Social_Context_rec2 =
      ifelse(
        str_detect(Social_Context_rec, "_"),
        Social_Context_rec,
        "with_known_others"        
      ),
    Social_Context_rec2 = factor(Social_Context_rec2)
    ) %>%
  dplyr::select(-Social_Context_rec)
```

### parsing dates

```{r parsing dates}
emo_mw_usable_clean <-
  emo_mw_usable_clean %>%
  mutate(
    emo_trigger_date = dmy(emo_trigger_date),
    emo_survey_start_date = dmy(emo_survey_start_date)
  )
```

### creating day and week variables

```{r}
library(lubridate)
# creating day and week variables
emo_mw_usable_clean_day <-
  emo_mw_usable_clean %>%
  mutate(
    day = day(emo_trigger_date),
    wday = lubridate::wday(emo_trigger_date, label = TRUE, abbr = TRUE),
    ELS_ID = factor(ELS_ID)
  ) 

# day num
emo_mw_usable_clean_dayorder  <-
  emo_mw_usable_clean_day %>%
  dplyr::group_by(ELS_ID) 

emo_mw_usable_clean_dayorder<- emo_mw_usable_clean_dayorder[order(emo_mw_usable_clean_dayorder$ELS_ID), ]
```

### wkday or wkend

```{r}
emo_mw_usable_clean_dayorder2 <-
  emo_mw_usable_clean_dayorder %>%
  mutate(
    week =
      ifelse(
        wday == "Sun" |
          wday == "Sat",
        "wkend",
        "wkday"
      )
  ) 
```


## merging with  summary sheet
```{r}
emo_merge <-
  left_join(
    emo_mw_usable_clean_dayorder2,
    sum_sheet_df,
    by = "ELS_ID"
  )

emo_merge2 <-
  emo_merge %>%
  filter(simul == "1")

emo_merge3 <-
  emo_merge2 %>%
  mutate(
    use_ema_date_t3 = mdy(use_ema_date_t3),
    use_ema_date_t4 = mdy(use_ema_date_t4),
    num_of_days_btwn_t3 =
      abs((use_ema_date_t3 %--% emo_survey_start_date)/ddays(1)),
    num_of_days_btwn_t4 =
      abs((use_ema_date_t4 %--% emo_survey_start_date)/ddays(1)),
    num_of_days_btwn =
      ifelse(
        is.na(num_of_days_btwn_t3),
        num_of_days_btwn_t4,
        num_of_days_btwn_t3
      )
    ) %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::mutate(
    startdate = min(emo_trigger_date),
    dayorder = abs((emo_trigger_date %--% startdate)/ddays(1))
  ) %>%
  relocate(startdate, .after = emo_trigger_date) %>%
  relocate(dayorder, .after = startdate) %>%
  mutate(
    timept = ifelse(!is.na(simiul_ahc_t3), "T3", "T4")
  )

emo_merge4 <-
  emo_merge3 %>%
  filter(
    dayorder < 15
  ) %>% # now max is 14 days
  dplyr::select(-c(num_of_days_btwn_t3, num_of_days_btwn_t4, starts_with("use"), simiul_ahc_t3, simul_ahc_t4))

emo_merge5 <-
  emo_merge4 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::mutate(
    emo_num_resp = n(),
    emo_rr = emo_num_resp/42
  )

```

### summmary

```{r}
emo_merge5 %>%
  dplyr::group_by(ELS_ID) %>%
  dplyr::summarise(
    n = n()
  ) # 106 subjects

write_csv(emo_merge5, "simul_ema_EMO_subjs.csv")
```

# Data viz

> for those with simulataneous ema and actigraphy, looking at happy and sad just as an example of what we would expect 


```{r viz sleep sat and emo}
emo_merge5 %>%
  ggplot(
    aes(dayorder, scale(Daily_Sad), group = ELS_ID)
    ) +
  geom_line(color = "blue") +
  geom_line(aes(x = dayorder, y = scale(Daily_Happy), group = ELS_ID), color = "pink") +
  facet_wrap(~ELS_ID) +
  scale_x_continuous(limits = c(1,15)) +
  theme_classic() +
  theme(
  strip.text.x = element_blank(),
  axis.title.y = element_text(size = 10)
  ) +
  labs(
    x = "Day",
    y = "Sleep Satisfaction (dark blue), Daily Sadness (pink)",
    title = "Trajectories of Self-Reported Sadness and Happiness"
    )

ggsave("daily_mw_emo_example_traj.png", width = 7, height = 5.5, dpi = 400)
```


### distribution of weekday vs. weekend and association with sadness and happiness

```{r}
level_key <- c(wkday = "wkday", wkend = "wkend")
emo_merge6 <-
  emo_merge5 %>%
  mutate(
    week = factor(week),
    week =  recode_factor(week,!!!level_key),
    dayorder = as.numeric(dayorder)
  )

# summary
emo_merge6 %>%
  dplyr::group_by(ELS_ID, week) %>%
  dplyr::summarise(
    n = n()
  )


emo_merge6 %>%
  drop_na(Daily_Sad) %>%
  drop_na(Daily_Happy) %>%
  drop_na(week) %>%
  ggplot(
    aes(
      x = dayorder, 
      y = scale(Daily_Sad)
    )
  ) +
  geom_line(aes(group = ELS_ID), color = "blue", alpha = .5) +
  geom_smooth(se=FALSE, size = 2, colour = "darkblue") +
  geom_line(aes(x = dayorder, y = scale(Daily_Happy), group = ELS_ID), color = "darkgoldenrod2", alpha = .5) +
  geom_smooth(aes(x = dayorder, y = scale(Daily_Happy)), size = 2, se = FALSE, colour = "darkgoldenrod2") +
  facet_wrap(~week) +
  theme_classic() +
  scale_x_continuous(
    limits = c(1, 14),                    
    breaks = seq(1, 14, 1)
    ) +
  labs(
    x = "Day",
    y = "Sad = blue, Happy= gold"
  )
ggsave("sleepsat_and_hours_byDay_byWkdayWkend.png", width = 7, height = 5)

```

#### tests
```{r}
# sad level differences by weekday/wkend 
summary(lmer(scale(Daily_Sad) ~ scale(dayorder) +factor(week) + (1 + scale(dayorder)|ELS_ID), data = emo_merge6))
## no effect of weekday/weekend

# sleep duration differences by weekday/wkend 
summary(lmer(scale(Daily_Happy) ~ scale(dayorder) + factor(week) + (1 + scale(dayorder)|ELS_ID), data = emo_merge6))
##  happier on weekends compared to weekdays
```

### write

```{r}
write_csv(emo_merge6, "~/Dropbox/JSK_SC/_scripts/2_at-home_wrangling/mw_daily_emo_longform.csv")
```

### timept data
```{r}
emo_merge6_red <-
  emo_merge6 %>%
  distinct(ELS_ID)
```


