---
title: "AHC_daily_ema_sleep_and_puberty"
author: "Jackie Kirshenbaum"
date: "7/25/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

>  This script analyzes adolescents only in grade school (before transition to college)

# Load Libraries
```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(foreign)
library(readxl)
library(haven) 
library(expss)
library(chron)
library(hms)
library(lubridate)
library(lmerTest)
```

# Reading in ema, actigraphy, and puberty/demo dfs

```{r,message=FALSE, filepaths}

# full AHC sample with pubertal and demographic info
pub_demo_act_ema_comb_fp <- "~/Dropbox/JSK_SC/1_sleep_puberty/_scripts_revised/3_aim1/pub_ema_act_merged.csv"

ahc_subj_df <-
  read_csv(pub_demo_act_ema_comb_fp) %>%
  dplyr::mutate(ELS_ID = factor(ELS_ID)) %>%
  as_tibble() %>%
  dplyr::select(-contains("act"))

```

# Load Libraries
```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(foreign)
library(readxl)
library(haven) 
library(expss)
library(chron)
library(hms)
library(lubridate)
library(lmerTest)
```


```{r}
options(scipen = 999, digits = 5)
```

# Testing association between pubertal stage and subjective sleep quality

## removing outliers in my DV 
```{r}
# outliers: outside 1.5 times the interquartile range above the upper quartile and below the lower quartile (Q1 - 1.5 * IQR or Q3 + 1.5 * IQR).
boxplot.stats(ahc_subj_df$dailysleep_sat_mean)$out # 0 outliers

write_csv(ahc_subj_df, "ahc_full_samp.csv")
```

## centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_subj_cent <-
  ahc_subj_df %>%
  dplyr::filter(high_school == "1") %>% 
  dplyr::mutate(
    Sex = factor(sex),
    COVID = factor(COVID),
    mw_dailysleep_sat_z = as.numeric(scale(dailysleep_sat_mean)),
    mw_dailysleepshrs_z = as.numeric(scale(dailysleep_hrs_mean)),
    cmep_z = as.numeric(scale(cmep_total)),
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    sleep_rr_z = as.numeric(scale(sleep_rr))
    ) %>%
  dplyr::group_by(Sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    tanner_average_t1_z = as.numeric(scale(tanner_average.T1)),
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc)),
    Age.T1_z = as.numeric(scale(Age.T1))
  ) %>%
  dplyr::ungroup()

# creating relative pubertal stage
rel_pub_mod <- lm(scale(tanner_average_ahc_z) ~ scale(Age_ahc_z), data = ahc_subj_cent)
summary(rel_pub_mod)
ahc_subj_cent <-
  ahc_subj_cent %>% 
  add_residuals(rel_pub_mod, var = "rel_pub_stg")

```

## Average RR

```{r}
summary(ahc_subj_cent$sleep_rr)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  #  0.357   0.500   0.786   0.726   0.929   1.000 
```
### NAs for cmep and stress

```{r}
sum(is.na(ahc_subj_cent$cmep_total)) # 3
sum(is.na(ahc_subj_cent$ASQ_total)) # 1
```

## Testing potential covariates: asociations with sleep satisfaction/restfulness

```{r}
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$BMI_ahc) # not sig
# r=-0.052372, p=0.68
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$Household_Income_ahc) #not sig
# r=0.16577, p=0.17
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$sleep_rr) # not sig
# r=0.062011, p=0.6
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$sumsev_type_t1) # not sig
# r=-0.17514, p=0.14
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$ASQ_total) # sig
# r=-0.32728, p=0.0047
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$cmep_total) # sig
# r=0.33684, p=0.0041
summary(lm(mw_dailysleep_sat_z ~ Sex, data = ahc_subj_cent)) # not sig 
# B=-0.214, p=0.39 
summary(lm(mw_dailysleep_sat_z ~ COVID, data = ahc_subj_cent)) # not sig 
# B=-0.0888, p=0.71 
```

## extra libraries
```{r,message=FALSE,echo=FALSE}
library(gvlma)
library(car)
library(BayesFactor)
library(bayestestR)
library(sjPlot)
```


## Association between pubertal stage score and sleep sat

```{r}
# tanner stage
pubsleep <- lm(mw_dailysleep_sat_z ~ tanner_average_ahc_z, data = ahc_subj_cent)
summary(gvlma(pubsleep))
summary(pubsleep)
pubsleepplot <- gvlma(pubsleep)
plot.gvlma(pubsleepplot)


# negative association between tanner stage and sleep sat
# (B=-0.326334393457243466, p=0.0048)
pubsleep_int <- summary(pubsleep)$coefficients[1,1]
pubsleep_slp <- summary(pubsleep)$coefficients[2,1]
# CI
ci_pubsleep = Boot(pubsleep, coef,
method='case', R=1000)
boot_confint_ci_pubsleep <- confint(ci_pubsleep, type='norm')
#                       2.5 %    97.5 %
# tanner_average_ahc_z -0.52504 -0.11717


# Bayes Factor Tanner Stage
bayes_pub <- regressionBF(
  formula = mw_dailysleep_sat_z ~ tanner_average_ahc_z, data = ahc_subj_cent
) # 8.1767 

```

### Repeating with Age

```{r}
agesleep <- lm(mw_dailysleep_sat_z ~ Age_ahc_z, data = ahc_subj_cent)
summary(gvlma(agesleep))
summary(agesleep)
agesleepplot <- gvlma(agesleep)
plot.gvlma(agesleepplot)

# no association btwn age and sleep
# (B=-0.1409, p=0.24)

agesleep_int <- summary(agesleep)$coefficients[1,1]
agesleep_slp <- summary(agesleep)$coefficients[2,1]
# CI
ci_agesleep = Boot(agesleep, coef,
method='case', R=1000)
boot_confint_ci_agesleep <- confint(ci_agesleep, type='norm')
#               2.5 %    97.5 %
# Age_ahc_z    -0.35624 0.11227

# Bayes Factor
bayes_age <- regressionBF(
  formula = mw_dailysleep_sat_z ~ Age_ahc_z, data = ahc_subj_cent
) # 0.39582 

```
### Plotting Tanner vs Age
```{r}
# plotting
colors <- c("Tanner stage" = "blue", "Age" = "red")
p1 <-
  ahc_subj_cent %>%
  ggplot(
    aes(x=tanner_average_ahc_z, y=mw_dailysleep_sat_z, color = "Tanner stage")
  ) +
  geom_point(size = 2, alpha = .5) +
  geom_point(
    aes(x = Age_ahc_z, y = mw_dailysleep_sat_z, color = "Age"), size = 2, alpha = .5
  ) +
  geom_abline(
    intercept = agesleep_int,
    slope = agesleep_slp,
    size = 2,
    color = "red"
  ) +
  geom_abline(
    intercept = pubsleep_int, 
    slope = pubsleep_slp,
    size=2,
    color = "blue"
    ) +
  labs(
    x = "Development Stage (z-scored)",
    y = "Averaged Daily Sleep Satisfaction (z-scored)",
    title = "Developmental Stage and Averaged Daily Sleep Satisfaction",
    color = "Developmental Stage"
  ) +
  scale_color_manual(values = colors) +
  theme_classic() +
  theme(plot.title = element_text(size = 9),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9))
p1
ggsave("dev_stage_daily_sleep_sat.tiff", p1, width = 6, height = 5)
```

### Controlling for perceived stress severity and circadian preference
```{r}
ahc_cent_bf <-
  ahc_subj_cent %>%
  drop_na(ASQ_total) %>%
  drop_na(cmep_total)

pubasqcmep <- lm(mw_dailysleep_sat_z ~ tanner_average_ahc_z + cmep_z + ASQ_total_z, data = ahc_cent_bf)
summary(gvlma(pubasqcmep))
summary(pubasqcmep)
pubasqcmepplot <- gvlma(pubasqcmep)
plot.gvlma(pubasqcmepplot)
ci_pubasqcmep = Boot(pubasqcmep, coef,
method='case', R=1000)
ci_relpubasqcmepboot <- confint(ci_pubasqcmep, type='norm')
#                          2.5 %    97.5 %
# (Intercept)          -0.239401  0.20653516
# tanner_average_ahc_z -0.4142701 -0.001207
# cmep_z               -0.0021835  0.419004
# ASQ_total_z          -0.4145017  0.015578
regressionBF(
  formula = mw_dailysleep_sat_z ~ tanner_average_ahc_z + ASQ_total_z + cmep_z, data = ahc_cent_bf
) # tanner_average_ahc_z pub: 6.8191, asq: 7.2692 , cmep: 9.4864 

```


# Testing association between pubertal stage and subjective sleep duration

## removing outliers in my DV 
```{r}
library(rstatix)
library(moments)
boxplot.stats(ahc_subj_df$dailysleep_hrs_mean)$out # 1 outlier
durout <- boxplot.stats(ahc_subj_df$dailysleep_hrs_mean)$out
duroutid <-  which(ahc_subj_df$dailysleep_hrs_mean %in% c(durout))
ahc_subj_df[duroutid, ]$ELS_ID 

ahc_subj_df_rem <-
  ahc_subj_df %>%
  filter(
    !ELS_ID == "96"
  ) 

```


## centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_subj_df_rem_cent_hrs <-
  ahc_subj_df_rem %>%
  dplyr::mutate(
    Sex = factor(sex),
    mw_dailysleep_sat_z = as.numeric(scale(dailysleep_sat_mean)),
    mw_dailysleepshrs_z = as.numeric(scale(dailysleep_hrs_mean)),
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    cmep_z = as.numeric(scale(cmep_total)),    
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    sleep_rr_z = as.numeric(scale(sleep_rr))
    ) %>%
  dplyr::group_by(Sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    tanner_average_t1_z = as.numeric(scale(tanner_average.T1)),
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc)),
    Age.T1_z = as.numeric(scale(Age.T1))
  ) %>%
  ungroup()

```

## Testing potential covariates: asociations with sleep duration

```{r}
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$BMI_ahc_z) # not sig
# r=-0.098023, p=0.44
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$Household_Income_ahc_z) # not sig
# r=0.18155, p=0.13
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$sumsev_type_z) # sig
# r=-0.27706 , p=0.017
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$ASQ_total) # not sig
# r=-0.204, p=0.081
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$cmep_total) # sig
# r=0.22077, p=0.062
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$sleep_rr) # sig
# r = 0.16078, p=0.17
summary(lm(mw_dailysleepshrs_z ~ Sex, data = ahc_subj_df_rem_cent_hrs)) # marginal
# B=0.402, p=0.10 
summary(lm(mw_dailysleepshrs_z ~ COVID, data = ahc_subj_df_rem_cent_hrs)) # marginal
# B=0.381, p=0.11 
```

## Association between pubertal stage score and sleep dur

```{r}
# Tanner
pubsleepdur <- lm(mw_dailysleepshrs_z ~ tanner_average_ahc_z, data = ahc_subj_df_rem_cent_hrs)
summary(gvlma(pubsleepdur))
summary(pubsleepdur)
pubsleepdurplot <- gvlma(pubsleepdur)
plot.gvlma(pubsleepdurplot)

# Tempo
temposleepdur <- lm(mw_dailysleepshrs_z ~ tanner_average_t1_z, data = ahc_subj_df_rem_cent_hrs)
summary(gvlma(temposleepdur))

# negative association between relative pubertal stage and sleep dur
# (B=-0.114449697228927144, p=0.33)

# CI
ci_pubsleepdur = Boot(pubsleepdur, coef,
method='case', R=1000)
boot_confint_ci_pubsleepdur <- confint(ci_pubsleepdur, type='norm')
#                       2.5 %    97.5 %
# tanner_average_ahc_z -0.33389 0.10840

# Bayes Factor
bayes_pubsleepdur <- regressionBF(
  formula = mw_dailysleepshrs_z ~ tanner_average_ahc_z, data = ahc_subj_df_rem_cent_hrs
) # 0.36059 


```


##### LONG FORM ######

```{r}
# full AHC sample with pubertal and demographic info LONG FORM
pub_demo_act_ema_comb_long_fp <- "~/Dropbox/JSK_SC/1_sleep_puberty/_scripts_revised/3_aim1/pub_ema_act_merged_long.csv"

ahc_long_df <-
  read_csv(pub_demo_act_ema_comb_long_fp) %>%
  dplyr::mutate(
    ELS_ID = factor(ELS_ID),
    high_school = factor(high_school)
    ) %>%
  as_tibble()
```

#### centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_long_df_cent <-
  ahc_long_df %>%
  dplyr::filter(high_school == "1") %>% # results hold filtering for those less than 19
  dplyr::group_by(ELS_ID) %>%
  dplyr::mutate(
    Sex = factor(sex),
    mw_dailysleep_sat_z = as.numeric(scale(DailySleep_satisfaction)),
    mw_dailysleepshrs_z = as.numeric(scale(DailySleep_hrs_rec))
    ) %>%
  dplyr::group_by(Sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    cmep_z = as.numeric(scale(cmep_total)),
    sleep_rr_z = as.numeric(scale(sleep_rr)),
    respnum_z = as.numeric(scale(respnumber))
  )
```

#### Examining the effect of Tanner stage on daily sleep satisfaction in long form

```{r}

sleepsat_long <- 
  lmer(
    DailySleep_satisfaction ~ 
      tanner_average_ahc_z  + respnum_z + ASQ_total_z + cmep_z +
         (1 + respnum_z | ELS_ID), 
       data = ahc_long_df_cent
  )
summary(sleepsat_long)

sleepsat_long_age <- 
  lmer(
    DailySleep_satisfaction ~ 
      Age_ahc_z  + respnum_z + ASQ_total_z + cmep_z +
         (1 + respnum_z | ELS_ID), 
       data = ahc_long_df_cent
  )
summary(sleepsat_long_age)
```

# Summary

> Pubertal stage, but not age, is associated with daily sleep satisfaction. Covariate of circadian preference was also still associated with sleep satisfaction; adolescent stress was not associated with daily sleep satisfaction anymore (holds in long form too). Pubertal stage was not associated with daily sleep duration. 

