---
title: "AHC_daily_ema_sleep_and_puberty_supplementary"
author: "Jackie Kirshenbaum"
date: "4/12/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

> This script analyzes all participants in high school + post-highschool 

# Load Libraries
```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(foreign)
library(readxl)
library(haven) 
library(expss)
library(chron)
library(hms)
library(lubridate)
library(lmerTest)
```

# Reading in ema, actigraphy, and puberty/demo dfs

```{r,message=FALSE, filepaths}

# full AHC sample with pubertal and demographic info
pub_demo_act_ema_comb_fp <- "~/Dropbox/JSK_SC/_scripts/3_aim1/pub_ema_act_merged.csv"

ahc_subj_df <-
  read_csv(pub_demo_act_ema_comb_fp) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  as_tibble() %>%
  dplyr::select(-contains("act"))

```

# Load Libraries
```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(foreign)
library(readxl)
library(haven) 
library(expss)
library(chron)
library(hms)
library(lubridate)
library(lmerTest)
```


```{r}
options(scipen = 999, digits = 5)
```

# Testing association between pubertal stage and subjective sleep quality

## removing outliers in my DV 
```{r}
# outliers: outside 1.5 times the interquartile range above the upper quartile and below the lower quartile (Q1 - 1.5 * IQR or Q3 + 1.5 * IQR).
boxplot.stats(ahc_subj_df$dailysleep_sat_mean)$out # 0 outliers
```

## centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_subj_cent <-
  ahc_subj_df %>%
  mutate(
    Sex = factor(sex),
    COVID = factor(COVID),
    mw_dailysleep_sat_z = as.numeric(scale(dailysleep_sat_mean)),
    mw_dailysleepshrs_z = as.numeric(scale(dailysleep_hrs_mean)),
    cmep_z = as.numeric(scale(cmep_total)),
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    sleep_rr_z = as.numeric(scale(sleep_rr))
    ) %>%
  group_by(Sex) %>%
  mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc))
  ) %>%
  ungroup()
```

## Average RR

```{r}
summary(ahc_subj_cent$sleep_rr)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  #  0.214   0.429   0.714   0.665   0.929   1.000 
```
### NAs for cmep and stress

```{r}
sum(is.na(ahc_subj_cent$cmep_total)) # 5
sum(is.na(ahc_subj_cent$ASQ_total)) # 2
```

## Testing potential covariates: asociations with sleep satisfaction/restfulness

```{r}
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$BMI_ahc) # not sig
# r=-0.041526, p=0.72
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$Household_Income_ahc) #not sig
# r=0.13752, p=0.21
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$sleep_rr) # not sig
# r=0.012469, p=0.91
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$sumsev_type_t1) # not sig
# r=-0.10489, p=0.33
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$ASQ_total) # sig
# r=-0.28032, p=0.0085
cor.test(ahc_subj_cent$dailysleep_sat_mean, ahc_subj_cent$cmep_total) # sig
# r=0.37248, p=0.00048
summary(lm(mw_dailysleep_sat_z ~ Sex, data = ahc_subj_cent)) # not sig 
# B=-0.204, p=.36 
summary(lm(mw_dailysleep_sat_z ~ COVID, data = ahc_subj_cent)) # not sig 
# B=-0.0541, p=0.81 
```

## extra libraries
```{r,message=FALSE,echo=FALSE}
library(gvlma)
library(car)
library(BayesFactor)
library(bayestestR)
library(sjPlot)
```


## Association between pubertal stage score and sleep sat

```{r}
pubsleep <- lm(mw_dailysleep_sat_z ~ tanner_average_ahc_z, data = ahc_subj_cent)
summary(gvlma(pubsleep))
summary(pubsleep)
pubsleepplot <- gvlma(pubsleep)
plot.gvlma(pubsleepplot)

# negative association between relative pubertal stage and sleep sat
# (B=-0.276131635114858220, p=0.0092)

pubsleep_int <- summary(pubsleep)$coefficients[1,1]
pubsleep_slp <- summary(pubsleep)$coefficients[2,1]
# CI
ci_pubsleep = Boot(pubsleep, coef,
method='case', R=1000)
boot_confint_ci_pubsleep <- confint(ci_pubsleep, type='norm')
#                       2.5 %    97.5 %
# tanner_average_ahc_z -0.46490 -0.084348

# Bayes Factor
bayes_pub <- regressionBF(
  formula = mw_dailysleep_sat_z ~ tanner_average_ahc_z, data = ahc_subj_cent
) #  4.6799 


```

### Repeating with Age

```{r}
agesleep <- lm(mw_dailysleep_sat_z ~ Age_ahc_z, data = ahc_subj_cent)
summary(gvlma(agesleep))
summary(agesleep)
agesleepplot <- gvlma(agesleep)
plot.gvlma(agesleepplot)

# no association btwn age and sleep
# (B=-0.0935154500193696170, p=0.39)

agesleep_int <- summary(agesleep)$coefficients[1,1]
agesleep_slp <- summary(agesleep)$coefficients[2,1]
# CI
ci_agesleep = Boot(agesleep, coef,
method='case', R=1000)
boot_confint_ci_agesleep <- confint(ci_agesleep, type='norm')
#               2.5 %    97.5 %
# Age_ahc_z   -0.29517 0.13707

# Bayes Factor
bayes_age <- regressionBF(
  formula = mw_dailysleep_sat_z ~ Age_ahc_z, data = ahc_subj_cent
) # 0.30962 

```
### Plotting Tanner vs Age
```{r}
# plotting
colors <- c("Tanner stage" = "blue", "Age" = "red")
p1 <-
  ahc_subj_cent %>%
  ggplot(
    aes(x=tanner_average_ahc_z, y=mw_dailysleep_sat_z, color = "Tanner stage")
  ) +
  geom_point(size = 2, alpha = .5) +
  geom_point(
    aes(x = Age_ahc_z, y = mw_dailysleep_sat_z, color = "Age"), size = 2, alpha = .5
  ) +
  geom_abline(
    intercept = agesleep_int,
    slope = agesleep_slp,
    size = 2,
    color = "red"
  ) +
  geom_abline(
    intercept = pubsleep_int, 
    slope = pubsleep_slp,
    size=2,
    color = "blue"
    ) +
  labs(
    x = "Development Stage (z-scored)",
    y = "Averaged Daily Sleep Satisfaction (z-scored)",
    title = "Developmental Stage and Averaged Daily Sleep Satisfaction",
    color = "Developmental Stage"
  ) +
  scale_color_manual(values = colors) +
  theme_classic() +
  theme(plot.title = element_text(size = 9),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9))
p1
```

### Controlling for perceived stress severity and circadian preference
```{r}
ahc_cent_bf <-
  ahc_subj_cent %>%
  drop_na(ASQ_total) %>%
  drop_na(cmep_total)
pubasqcmep <- lm(mw_dailysleep_sat_z ~ tanner_average_ahc_z + ASQ_total_z + cmep_z, data = ahc_cent_bf)
summary(gvlma(pubasqcmep))
summary(pubasqcmep)
pubasqcmepplot <- gvlma(pubasqcmep)
plot.gvlma(pubasqcmepplot)
ci_pubasqcmep = Boot(pubasqcmep, coef,
method='case', R=1000)
ci_relpubasqcmepboot <- confint(ci_pubasqcmep, type='norm')
#                         2.5%      97.5%
# tanner_average_ahc_z -0.387826 -0.0029039
# ASQ_total_z          -0.339137  0.0681323
# cmep_z                0.095139  0.4844219
# Bayes Factor
regressionBF(
  formula = mw_dailysleep_sat_z ~ tanner_average_ahc_z + ASQ_total_z + cmep_z, data = ahc_cent_bf
) 
# tanner: 4.7623
# ASQ_total_z: 4.6372 ±0%
# cmep_z 56.152 ±0%

```


# Testing association between pubertal stage and subjective sleep duration

## removing outliers in my DV 
```{r}
library(rstatix)
library(moments)
boxplot.stats(ahc_subj_df$dailysleep_hrs_mean)$out # 1 outlier
durout <- boxplot.stats(ahc_subj_df$dailysleep_hrs_mean)$out
duroutid <-  which(ahc_subj_df$dailysleep_hrs_mean %in% c(durout))
ahc_subj_df[duroutid, ]$ELS_ID 

ahc_subj_df_rem <-
  ahc_subj_df %>%
  filter(
    !ELS_ID == "96"
  ) 

```


## centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_subj_df_rem_cent_hrs <-
  ahc_subj_df_rem %>%
  dplyr::mutate(
    Sex = factor(sex),
    mw_dailysleep_sat_z = as.numeric(scale(dailysleep_sat_mean)),
    mw_dailysleepshrs_z = as.numeric(scale(dailysleep_hrs_mean)),
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    cmep_z = as.numeric(scale(cmep_total)),    
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    sleep_rr_z = as.numeric(scale(sleep_rr))
    ) %>%
  dplyr::group_by(Sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc))
  ) %>%
  ungroup()
```

## Testing potential covariates: asociations with sleep duration

```{r}
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$BMI_ahc_z) # not sig
# r=-0.10727, p=0.35
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$Household_Income_ahc_z) # not sig
# r=0.17024, p=0.12
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$sumsev_type_z) # sig
# r=-0.28631 , p=0.0072
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$ASQ_total) # not sig
# r=-0.21032, p=0.052
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$cmep_total) # sig
# r=0.23832, p=0.03
cor.test(ahc_subj_df_rem_cent_hrs$mw_dailysleepshrs_z, ahc_subj_df_rem_cent_hrs$sleep_rr) # sig
# r = 0.24212, p=0.023
summary(lm(mw_dailysleepshrs_z ~ Sex, data = ahc_subj_df_rem_cent_hrs)) # marginal
# B=0.387, p=0.083 
summary(lm(mw_dailysleepshrs_z ~ COVID, data = ahc_subj_df_rem_cent_hrs)) # marginal
# B=0.397, p=0.073 
```

## Association between pubertal stage score and sleep dur

```{r}
pubsleepdur <- lm(mw_dailysleepshrs_z ~ tanner_average_ahc_z, data = ahc_subj_df_rem_cent_hrs)
summary(gvlma(pubsleepdur))
summary(pubsleepdur)
pubsleepdurplot <- gvlma(pubsleepdur)
plot.gvlma(pubsleepdurplot)

# negative association between relative pubertal stage and sleep dur
# (B=-0.106076474916116201, p=0.33)
# CI
ci_pubsleepdur = Boot(pubsleepdur, coef,
method='case', R=1000)
boot_confint_ci_pubsleepdur <- confint(ci_pubsleepdur, type='norm')
#                       2.5 %    97.5 %
# tanner_average_ahc_z -0.31940 0.10287

# Bayes Factor
bayes_pubsleepdur <- regressionBF(
  formula = mw_dailysleepshrs_z ~ tanner_average_ahc_z, data = ahc_subj_df_rem_cent_hrs
) # 0.34074 


```

# Summary

> Results are consistent with main analysis. Effect of pubertal stage in relation to daily sleep satisfaction becomes marginal when controlling for circadian pref and perceived stress severity. 

