---
title: "AHC_obj_act_sleep_and_puberty_supplementary"
author: "Jackie Kirshenbaum"
date: "4/12/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

> This script analyzes all participants in high school + post-highschool 


# Load Libraries
```{r, message=FALSE,echo=FALSE,warning=FALSE}
library(tidyverse)
library(modelr)
library(foreign)
library(readxl)
library(haven) 
library(expss)
library(chron)
library(hms)
library(lubridate)
library(lmerTest)
```

# Reading in ema, actigraphy, and puberty/demo dfs

```{r,message=FALSE, filepaths}

# full AHC sample with pubertal and demographic info
pub_demo_act_ema_comb_fp <- "~/Dropbox/JSK_SC/1_sleep_puberty/_scripts/3_aim1/pub_ema_act_merged.csv"

ahc_obj_df <-
  read_csv(pub_demo_act_ema_comb_fp) %>%
  mutate(ELS_ID = factor(ELS_ID)) %>%
  as_tibble() %>%
  dplyr::select(-contains("daily"), -sleep_num_resp, -sleep_rr)

```

# Load Libraries
```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(foreign)
library(readxl)
library(haven) 
library(expss)
library(chron)
library(hms)
library(lubridate)
library(lmerTest)
```


```{r}
options(scipen = 999, digits = 5)
```

# Testing association between pubertal stage and objective sleep efficiency

## removing outliers in my DV 
```{r}
# outliers: outside 1.5 times the interquartile range above the upper quartile and below the lower quartile (Q1 - 1.5 * IQR or Q3 + 1.5 * IQR).
boxplot.stats(ahc_obj_df$efficiency_mean)$out # 0 outliers
effout <- boxplot.stats(ahc_obj_df$efficiency_mean)$out
effoutid <-  which(ahc_obj_df$efficiency_mean %in% c(effout))
ahc_obj_df[effoutid, ]$ELS_ID 

ahc_obj_df_rem <-
  ahc_obj_df %>%
  filter(
    !ELS_ID == "143"
  ) 
```

## centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_obj_cent <-
  ahc_obj_df_rem %>%
  dplyr::mutate(
    Sex = factor(sex),
    COVID = factor(COVID),
    act_efficiency_z = as.numeric(scale(efficiency_mean)),
    act_sleepdur_z = as.numeric(scale(sleep_time_hrs_mean)),
    cmep_z = as.numeric(scale(cmep_total)),
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    act_rr_z = as.numeric(scale(act_rr))
    ) %>%
  dplyr::group_by(Sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc))
  ) %>%
  dplyr::ungroup()
```

## Average RR

```{r}
summary(ahc_obj_cent$act_rr)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#   0.214   0.857   1.000   0.994   1.071   2.786 
```


## Testing potential covariates: asociations with sleep efficiency

```{r}
contrasts(ahc_obj_cent$Sex) <- contr.treatment(2)
cor.test(ahc_obj_cent$efficiency_mean, ahc_obj_cent$BMI_ahc) # not sig
# r=0.026997, p=0.82
cor.test(ahc_obj_cent$efficiency_mean, ahc_obj_cent$Household_Income_ahc) #not sig
# r=0.092118, p=0.40
cor.test(ahc_obj_cent$efficiency_mean, ahc_obj_cent$act_rr) # not sig
# r=-0.0046555, p=0.97
cor.test(ahc_obj_cent$efficiency_mean, ahc_obj_cent$sumsev_type_t1) # not sig
# r=0.020487, p=0.85
cor.test(ahc_obj_cent$efficiency_mean, ahc_obj_cent$ASQ_total) # not sig
# r=0.15793, p=0.15
cor.test(ahc_obj_cent$efficiency_mean, ahc_obj_cent$cmep_total) # not sig
# r=-0.1886, p=0.088
summary(lm(scale(efficiency_mean) ~ Sex, data = ahc_obj_cent)) # not sig 
# B=0.268, p=0.23
summary(lm(scale(efficiency_mean) ~ COVID, data = ahc_obj_cent)) # not sig 
# B=0.1592, p=0.48
```

## extra libraries
```{r,message=FALSE}
library(gvlma)
library(car)
library(BayesFactor)
library(bayestestR)
library(sjPlot)
```


## Association between Tanner Stage and sleep efficiency

```{r}
effpub <- lm(act_efficiency_z ~ tanner_average_ahc_z, data = ahc_obj_cent)
summary(gvlma(effpub))
summary(effpub)
plot.gvlma(gvlma(effpub))

# CI
ci_effpub = Boot(effpub, coef,
method='case', R=1000)
boot_confint_pub <- confint(ci_effpub, type='norm')
#                         2.5 %    97.5 %
# tanner_average_ahc_z  -0.042397 0.37780

# Bayes Factor
bayes_pub <- regressionBF(
  formula = act_efficiency_z ~ tanner_average_ahc_z, data = ahc_obj_cent
) # 0.78559 


```
# Repeating with Age
```{r}
effage <- lm(act_efficiency_z ~ Age_ahc_z, data = ahc_obj_cent)
summary(gvlma(effage))
summary(effage)
plot.gvlma(gvlma(effage))

# CI
ci_effage = Boot(effage, coef,
method='case', R=1000)
boot_confint_age <- confint(ci_effage, type='norm')
#               2.5 %    97.5 %
# Age_ahc_z   -0.10083 0.30378

# Bayes Factor
bayes_age<- regressionBF(
  formula = act_efficiency_z ~ Age_ahc_z, data = ahc_obj_cent
) # 0.33295 


```

# Testing association between pubertal stage and objective sleep duration

## removing outliers in my DV 
```{r}
library(rstatix)
library(moments)
boxplot.stats(ahc_obj_df$sleep_time_hrs_mean)$out # 1 outlier
durout <- boxplot.stats(ahc_obj_df$sleep_time_hrs_mean)$out
duroutid <-  which(ahc_obj_df$sleep_time_hrs_mean %in% c(durout))
ahc_obj_df[duroutid, ]$ELS_ID 

ahc_obj_df_rem <-
  ahc_obj_df %>%
  filter(
    !ELS_ID == "85"
  ) 

```


## centering based on this sample
I standardized tanner stage, age at tanner assessment, and bmi within each sex
```{r}
ahc_obj_df_rem_cent_hrs <-
  ahc_obj_df_rem %>%
  dplyr::mutate(
    Sex = factor(sex),
    act_efficiency_z = as.numeric(scale(efficiency_mean)),
    act_sleepdur_z = as.numeric(scale(sleep_time_hrs_mean)),
    Household_Income_ahc_z = as.numeric(scale(Household_Income_ahc)),
    cmep_z = as.numeric(scale(cmep_total)),    
    sumsev_type_z = as.numeric(scale(sumsev_type_t1)),
    ASQ_total_z = as.numeric(scale(ASQ_total)),
    act_rr_z = as.numeric(scale(act_rr))
    ) %>%
  dplyr::group_by(Sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)), # standardize within each sex
    BMI_ahc_z = as.numeric(scale(BMI_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc))
  ) %>%
  dplyr::ungroup()

contrasts(ahc_obj_df_rem_cent_hrs$Sex) <- contr.treatment(2)

```



## Testing potential covariates: asociations with sleep duration

```{r}
cor.test(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean, ahc_obj_df_rem_cent_hrs$BMI_ahc) # not sig
# r=-0.036237, p=0.750
cor.test(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean, ahc_obj_df_rem_cent_hrs$Household_Income_ahc) # not sig
# r=0.16821, p=0.13
cor.test(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean, ahc_obj_df_rem_cent_hrs$sumsev_type_t1) # not sig
# r=-0.11585, p=0.29
cor.test(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean, ahc_obj_df_rem_cent_hrs$ASQ_total) # not sig
# r=0.086793, p=0.43
cor.test(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean, ahc_obj_df_rem_cent_hrs$cmep_total) # sig
# r=-0.22245, p=0.043
cor.test(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean, ahc_obj_df_rem_cent_hrs$act_rr) # not sig
# r=-0.099721, p=0.36
summary(lm(act_sleepdur_z ~ Sex, data = ahc_obj_df_rem_cent_hrs)) # sig
# B=0.837, p=0.0001

t.test(sleep_time_hrs_mean ~ Sex, data = ahc_obj_df_rem_cent_hrs)
# M for boys = 6.4300, M for girls = 7.0766
sd(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean[ahc_obj_df_rem_cent_hrs$Sex == "1"]) # male sd = 0.7326
sd(ahc_obj_df_rem_cent_hrs$sleep_time_hrs_mean[ahc_obj_df_rem_cent_hrs$Sex == "2"]) # female sd = 0.70037

summary(lm(act_sleepdur_z ~ COVID, data = ahc_obj_df_rem_cent_hrs)) # not sig
# B=0.0752, p=0.74
```


## Association between Tanner stage  and objective sleep duration

```{r}

pubhrs <- lm(act_sleepdur_z ~ tanner_average_ahc_z, data = ahc_obj_df_rem_cent_hrs)
summary(gvlma(pubhrs))
plot.gvlma(gvlma(pubhrs))

# CI
ci_pubhrs= Boot(pubhrs, coef,
method='case', R=1000)
boot_confint_ci_pubhrs <- confint(ci_pubhrs, type='norm')
#                           2.5%  97.5%
# tanner_average_ahc_z  -0.24239 0.2354

# Bayes Factor
bayes_pubhrs <- regressionBF(
  formula = act_sleepdur_z ~ tanner_average_ahc_z, data = ahc_obj_df_rem_cent_hrs
)
# 0.22295


```

## Association between Age  and objective sleep duration

```{r}

agehrs <- lm(act_sleepdur_z ~ Age_ahc_z, data = ahc_obj_df_rem_cent_hrs)
summary(gvlma(agehrs))
plot.gvlma(gvlma(agehrs))

# CI
ci_agehrs= Boot(agehrs, coef,
method='case', R=1000)
boot_confint_ci_agehrs <- confint(ci_agehrs, type='norm')
#               2.5%    97.5%
# Age_ahc_z   -0.29858 0.08334

# Bayes Factor
bayes_agehrs <- regressionBF(
  formula = act_sleepdur_z ~ Age_ahc_z, data = ahc_obj_df_rem_cent_hrs
)
# 0.3552


```


# Summary

> Neither Pubertal stage nor age is  associated with daily sleep efficiency or duration. 
