---
title: "3_ema_MLM_affectoutcome"
author: "Jackie"
date: "5/9/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

> This script analyzes adolescents only in grade school (before transition to college)


# Libraries

```{r, message=FALSE,warning=FALSE,echo=FALSE}
library(tidyverse)
library(modelr)
library(chron)
library(hms)
library(lubridate)
library(bootnet)
library(EMAtools)
library(lme4)
library(sjPlot)
library(sjstats)
library(assertthat)
library(effectsize)
library(tidyverse)
library(dplyr)
library(naniar)
library(data.table)
library(entropy)
library(psych)
```

## functions
```{r}
getmode <- function(v, na.rm=TRUE) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
# function from ben bolker
mylag <- function(x,lag) {
  c(rep(NA,lag),head(x,-lag))
}
```

# Read in data


```{r, message=FALSE}
emadffp <- "~/Dropbox/JSK_SC/_scripts/4_aim2/ema_and_act_merge.csv"
emadf <- read_csv(emadffp) %>%
  dplyr::filter(high_school == "1") %>%
  dplyr::select(
    ELS_ID, sex, triggerdate, dayorder, respnumber, trigger_tod, trigger_tod_order, wday, week, tanner_average_ahc,
    DailySleep_satisfaction:pos_total, emo_rr, sleep_rr, act_rr, Afternoon:COVID, device_OS
    ) %>%
  dplyr::mutate(
    ELS_ID = factor(ELS_ID),
    sex = factor(sex)
    )

emadf %>% distinct(ELS_ID)  

```

### filter out if less than 3 prompts of evening/afternoon/morning

```{r}
emadf_filt <-
  emadf %>%
  dplyr::group_by(ELS_ID, trigger_tod) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(
    toofew = ifelse(n < 3, "drop", "keep")
  )

emadf_filt %>%
  ggplot(
    aes(x = n, fill = trigger_tod)
  ) +
  facet_wrap(~trigger_tod) +
  geom_histogram(stat = "count")

emadf_filt_merge <-
  emadf_filt %>%
  dplyr::select(-trigger_tod, -n) %>%
  distinct() # 201 and 206 not enough data
  
```

```{r}
merge_restruct <-
  emadf %>% 
  filter(ELS_ID != "201") %>%
  filter(ELS_ID != "206") %>%
  ungroup() %>%
  dplyr::mutate(
    sex = factor(sex),
    COVID = factor(COVID),
    dayorder_f = factor(dayorder),
    tod_f = factor(trigger_tod_order)
  ) %>%
  dplyr::group_by(sex) %>%
  dplyr::mutate(
    tanner_average_ahc_z = as.numeric(scale(tanner_average_ahc)),
    Age_ahc_z = as.numeric(scale(Age_ahc)),
    BMI_ahc_z = as.numeric(scale(BMI_ahc))
  )
```


```{r}
merge_restruct <- merge_restruct[order(merge_restruct$ELS_ID), ]

```

# Centering

Level 1 variable centering
- computing deviations from the person's mean of that variable (x - mean(x)) for that person
Level 2 variable centering
- the person's mean (mean(x))

### First, we split our predictor variables into time-varying and time-invariant components

```{r}
library(plyr)
library(dplyr)
library(reshape2)
# daily sleep satisfaction
day.means_DailySleep_sat <- ddply(merge_restruct, "ELS_ID", summarize, DailySleep_sat.L2 =mean(DailySleep_satisfaction, na.rm=TRUE))
# daily sleep hours
day.means_DailySleep_hrs <- ddply(merge_restruct, "ELS_ID", summarize, DailySleep_hrs.L2 =mean(DailySleep_hrs_rec, na.rm=TRUE))
# efficiency
day.means_efficiency <- ddply(merge_restruct, "ELS_ID", summarize, efficiency.L2 =mean(efficiency, na.rm=TRUE))
# sleep hours
day.means_sleephrs <- ddply(merge_restruct, "ELS_ID", summarize, sleep_time.L2 =mean(sleep_time_hrs, na.rm=TRUE))
# neg_total
day.means_negtotal <- ddply(merge_restruct, "ELS_ID", summarize, NA.L2 =mean(neg_total, na.rm=TRUE))
# neg_total_sqrt
day.means_negtotal_sqrt <- ddply(merge_restruct, "ELS_ID", summarize, NAsqrt.L2 =mean(neg_total_sqrt, na.rm=TRUE))
# pos_total
day.means_postotal <- ddply(merge_restruct, "ELS_ID", summarize, PA.L2 =mean(pos_total, na.rm=TRUE))

#Calculating sample-centered versions *Note that this is done in a person-level data file.
day.means_DailySleep_sat$DailySleep_sat.L2.c <- as.numeric(scale(day.means_DailySleep_sat$DailySleep_sat.L2,center=TRUE,scale=FALSE))
day.means_DailySleep_hrs$DailySleep_hrs.L2.c <- as.numeric(scale(day.means_DailySleep_hrs$DailySleep_hrs.L2,center=TRUE,scale=FALSE))
day.means_efficiency$efficiency.L2.c <- as.numeric(scale(day.means_efficiency$efficiency.L2,center=TRUE,scale=FALSE))
day.means_sleephrs$sleep_time.L2.c <- as.numeric(scale(day.means_sleephrs$sleep_time.L2,center=TRUE,scale=FALSE))
day.means_negtotal$NA.L2.c <- as.numeric(scale(day.means_negtotal$NA.L2,center=TRUE,scale=FALSE))
day.means_negtotal_sqrt$NAsqrt.L2.c <- as.numeric(scale(day.means_negtotal_sqrt$NAsqrt.L2,center=TRUE,scale=FALSE))
day.means_postotal$PA.L2.c <- as.numeric(scale(day.means_postotal$PA.L2,center=TRUE,scale=FALSE))

#merging "trait" scores back into the *long* data file and calculate "state" scores.
day.means2 <- merge(
  merge_restruct,
  day.means_DailySleep_sat,
  by="ELS_ID"
  )
day.means3 <- merge(
  day.means2,
  day.means_DailySleep_hrs,
  by="ELS_ID"
  )
day.means4 <- merge(
  day.means3,
  day.means_efficiency,
  by="ELS_ID"
  )
day.means5 <- merge(
  day.means4,
  day.means_sleephrs,
  by="ELS_ID"
  )
day.means6 <- merge(
  day.means5,
  day.means_negtotal,
  by="ELS_ID"
  )
day.means7 <- merge(
  day.means6,
  day.means_negtotal_sqrt,
  by="ELS_ID"
  )
day.means8 <- merge(
  day.means7,
  day.means_postotal,
  by="ELS_ID"
  )



# creating level 1 variables
day.means8$DailySleep_sat.L1 <- day.means8$DailySleep_satisfaction - day.means8$DailySleep_sat.L2
day.means8$DailySleep_hrs.L1 <- day.means8$DailySleep_hrs_rec - day.means8$DailySleep_hrs.L2
day.means8$efficiency.L1 <- day.means8$efficiency - day.means8$efficiency.L2
day.means8$sleep_time.L1 <- day.means8$sleep_time_hrs - day.means8$sleep_time.L2
day.means8$NA.L1 <- day.means8$neg_total - day.means8$NA.L2
day.means8$NAsqrt.L1 <- day.means8$neg_total_sqrt - day.means8$NAsqrt.L2
day.means8$PA.L1 <- day.means8$pos_total - day.means8$PA.L2


# Centering time
day.means9 <-
  day.means8 %>%
  mutate(
    tod.c = scale(trigger_tod_order, center = TRUE, scale = FALSE),
    dayorder.c = scale(dayorder, center = TRUE, scale = FALSE)
  ) %>%
  relocate(tod.c, .after = trigger_tod_order) %>%
  relocate(dayorder.c, .after = dayorder)

```


# 3-level MLM to predict negative affect
> L1 = time of day (trigger_tod_order)
> L2 = sleep.state
> L3 = sleep.trait
> DV = NA
> _i_ = time of day
> _j_ = day
> _k_ = person

```{r,message=FALSE,echo=FALSE,warning=FALSE}
library(sjPlot)
library(lmerTest)
library(performance)
library(reghelper)
library(parameters)
library(MuMIn)
library(interactions)
```


### L1
\[NA_{ijk} = \beta_{0jk}+\beta_{1jk}TOD_{ijk}+\beta_{2jk}PA.State_{ijk}+e_{ijk}\]
```{r}

library(bruceR)
mlm_l1_na_ri <-
  lmer(
    neg_total_sqrt ~ 1 + trigger_tod_order + (1 | ELS_ID), data = day.means9
    )
mlm_l1_pa_ri <-
  lmer(
    pos_total ~ trigger_tod_order + (1 | ELS_ID), data = day.means9
    )


mlm_l1_na_rs <-
  lmer(
    neg_total_sqrt ~  1 + trigger_tod_order + (1 | ELS_ID/dayorder), data = day.means9)  # too complex with trigger_tod_order as a random slope
mlm_l1_pa_rs <-
  lmer(
    pos_total ~  1 + trigger_tod_order + (1 + trigger_tod_order | ELS_ID/dayorder), data = day.means9
    )

# model comparison
anova(mlm_l1_pa_ri, mlm_l1_pa_rs) # random slope better
anova(mlm_l1_na_ri, mlm_l1_na_rs) # random slope better

# summary
HLM_summary(mlm_l1_na_rs)
HLM_summary(mlm_l1_pa_rs)

```

> For NA, there is a marginal effect of day; no effect of time of day
> For PA, there is no effect of day, but there is an effect of time of day (positive affect increases in the evening compared to the morning)

### L2
\[\beta_{0jk}=\gamma_{00k}+\gamma_{01k}DailySleep_{jk}+\gamma_{02k}Day_{jk} + r_{0jk}\]  
\[\beta_{1jk}=\gamma_{10k}+\gamma_{11k}DailySleep_{jk}+\gamma_{12k}Day_{jk} +r_{1jk}\] 
\[\beta_{2jk}=\gamma_{20k}+r_{2jk}\] 
withindayslope (withinday affect trajecotry) = expectedvalue of withindayslope for person +sleep
```{r}
# sleep state is level 2

mlm_l2_na_ri_subj_sleep <-
  lmer(
    neg_total_sqrt ~  1 + dayorder + DailySleep_sat.L1*trigger_tod_order + PA.L1*trigger_tod_order + (1 | ELS_ID/dayorder), data = day.means9
    )
mlm_l2_na_rs_subj_sleep <-
    lmer(
      neg_total_sqrt ~  1  + dayorder + DailySleep_sat.L1*trigger_tod_order + PA.L1*trigger_tod_order + (1 + trigger_tod_order | ELS_ID/dayorder), data = day.means9
    )
anova(mlm_l2_na_ri_subj_sleep, mlm_l2_na_rs_subj_sleep) # random slope better
HLM_summary(mlm_l2_na_rs_subj_sleep)


mlm_l2_pa_ri_subj_sleep <-
  lmer(
    pos_total ~  1 + dayorder + DailySleep_sat.L1*trigger_tod_order + NAsqrt.L1*trigger_tod_order + (1 | ELS_ID/dayorder), data = day.means9
    )
mlm_l2_pa_rs_subj_sleep <-
  lmer(
    pos_total ~  1 + dayorder + DailySleep_sat.L1*trigger_tod_order + NAsqrt.L1*trigger_tod_order + (1 + trigger_tod_order | ELS_ID/dayorder), data = day.means9
    )
anova(mlm_l2_pa_ri_subj_sleep, mlm_l2_pa_rs_subj_sleep) # random slope better

HLM_summary(mlm_l2_pa_rs_subj_sleep)

```


### L3 (time-of-day * Daily Sleep Sat * puberty + Trait Sleep Sat * puberty)
\[\gamma_{00k}=\pi_{000}+\pi_{001}Puberty_{k}+\pi_{002}SleepTrait_{k}+\pi_{003}SleepTrait_{k}*Puberty_{k}+\pi_{004}PA.Trait_{k}+\nu_{00k}\]
\[\gamma_{10k}=\pi_{100}+\pi_{101}Puberty_{k}+\pi_{102}SleepTrait_{k}+\pi_{103}SleepTrait_{k}*Puberty_{k}+\nu_{10k}\]
\[\gamma_{01k}=\pi_{010}+\pi_{011}Puberty_{k}+\pi_{012}SleepTrait_{k}+\pi_{013}SleepTrait_{k}*Puberty_{k}+\nu_{01k}\]
\[\gamma_{02k}=\pi_{020}+\pi_{021}Puberty_{k}+\pi_{022}SleepTrait_{k}+\pi_{023}SleepTrait_{k}*Puberty_{k}+\nu_{02k}\]
\[\gamma_{11k}=\pi_{110}+\pi_{111}Puberty_{k}+\pi_{112}SleepTrait_{k}+\pi_{113}SleepTrait_{k}*Puberty_{k}+\nu_{11k}\]
\[\gamma_{12k}=\pi_{120}+\pi_{121}Puberty_{k}+\pi_{122}SleepTrait_{k}+\pi_{123}SleepTrait_{k}*Puberty_{k}+\nu_{12k}\]
\[\gamma_{20k}=\pi_{200}+\nu_{20k}\]

```{r}
day.means9_red <-
  day.means9 %>%
  dplyr::select(
    ELS_ID, dayorder, trigger_tod_order, tanner_average_ahc_z, neg_total_sqrt, pos_total, DailySleep_sat.L1, DailySleep_sat.L2.c, PA.L1, PA.L2.c, NAsqrt.L1, NAsqrt.L2.c
  )
# Neg Affect 
mlm_l3_na_ri_subj_sleep_sat_pub <-
  lmer(
    neg_total_sqrt ~ dayorder + trigger_tod_order*DailySleep_sat.L1*tanner_average_ahc_z  + DailySleep_sat.L2.c*tanner_average_ahc_z  + PA.L1 + PA.L2.c + (1  | ELS_ID/dayorder), data = day.means9_red
    ) 
mlm_l3_na_rs_subj_sleep_sat_pub <-
  lmer(
    neg_total_sqrt ~ dayorder + trigger_tod_order*DailySleep_sat.L1*tanner_average_ahc_z + DailySleep_sat.L2.c*tanner_average_ahc_z + PA.L1 + PA.L2.c + (1 + trigger_tod_order | ELS_ID/dayorder), data = day.means9_red
    ) 

anova(mlm_l3_na_ri_subj_sleep_sat_pub, mlm_l3_na_rs_subj_sleep_sat_pub)
HLM_summary(mlm_l3_na_rs_subj_sleep_sat_pub)


# Pos Affect
mlm_l3_pa_ri_subj_sleep_sat_pub <-
  lmer(
    pos_total ~ dayorder + trigger_tod_order*DailySleep_sat.L1*tanner_average_ahc_z + DailySleep_sat.L2.c*tanner_average_ahc_z + NAsqrt.L1 + NAsqrt.L2.c + (1  | ELS_ID/dayorder), data = day.means9_red
    ) 
mlm_l3_pa_rs_subj_sleep_sat_pub <-
  lmer(
    pos_total ~ dayorder + trigger_tod_order*DailySleep_sat.L1*tanner_average_ahc_z + DailySleep_sat.L2.c*tanner_average_ahc_z + NAsqrt.L1 + NAsqrt.L2.c + (1 + trigger_tod_order | ELS_ID/dayorder), data = day.means9_red
    ) 

anova(mlm_l3_pa_ri_subj_sleep_sat_pub, mlm_l3_pa_rs_subj_sleep_sat_pub)
HLM_summary(mlm_l3_pa_rs_subj_sleep_sat_pub)

library(sjPlot)
library(sjstats)

tab_model(mlm_l3_na_rs_subj_sleep_sat_pub, mlm_l3_pa_rs_subj_sleep_sat_pub, show.std = TRUE,  dv.labels = c("NA", "PA"), string.std_ci = "CI 95%", string.std.p = "p-value", col.order = c("std.est", "std.ci", "std.p"), pred.labels = c("Intercept", "day", "time-of-day", "DailySleepSat.State", "TannerStage", "DailySleepSat.Trait.c",  "PA.State", "PA.Trait", "DailySleepSat.State*time-of-day", "TannerStage*time-of-day", "DailySleepSat.State*TannerStage", "DailySleepSat.Trait.c*TannerStage", "DailySleepSat.State*TannerStage*time-of-day", "NA.State", "NA.Trait"), file = "affect_sleepsat_mod.doc")
```

